\LARGE \bf Preparation of Papers for IEEE Sponsored Conferences \& Symposia*

### 2017-02-08 intro.tex
.

We further our exploration of trajectory segment based methods; and
explore an alternative approach to overcome the explosion in abstract
states. Instead of selectively refining the abstraction, we compute a
model of the black box system and use bounded model checking to find a
concrete counter-example in the model~\footnote{Due to modeling
errors, this might not be reproducible in the original black-box
system.}. We then use this counter-example to guide the search towards
a counter-example in the original system.

@@

Towards the end of the chapter, we discuss extensions to data driven
approaches (instead of simulator driven), where instead of a
simulator, a fixed set of data is provided as the behavioral
description of the system. Using a combination of relational modeling,
and program analysis we outline the future work for falsifying
properties of SDCS.

\section{Overview}
\label{sec:ovw}

We now briefly describe the approach. Given a numerical simulator
$\simulate$ and an initial abstraction defined by a quantization
function $\quant_\epsilon$ and a time step $\Delta$, we use
scatter-and-simulate (parameterized by number of samples $n$ as
described in \chapref{abs}) to explore the abstract graph
$\scr{H}(\Delta)$. The result is a graph $G$, which has a finite
number of abstract states $C$ (or cells) and edges $(C,C')$ iff $C
\areach{\Delta} C'$.

Instead of using a CEGAR like loop (as in \chapref{abs}), we use the
generated trajectory segments to learn quantitative models describing
the local behavior of the system. These models are defined by a set of
relations $R \subseteq \HybridStates \times \HybridStates$ for each
edge of the reachability graph.

\input{figs/segtraj_rel_enriched.tex}

Recall that each edge $(C,C')$ of the graph $G$ denotes an observed
trajectory segment between the respective cells. The graph abstraction
$G$ only states that there exists a state $\x \in C$ from which the
system can evolve to a future state $\x' \in C'$. To increase the
precision we iteratively refined the abstraction by state
splitting. Instead, we now propose an `enrichment' $G^R$ of $G$ by computing a
set of local relations $R_{(C,C')}(\x,\x')$ for every edge $(C,C')$,
which non-deterministically describe relations between $\x \in C$ and $\x'
\in C'$. This is illustrated in \figref{enriched-edge} (compare with
\figref{segtraj}).

The enriched graph $G^R$ captures the underlying local forward
dynamics describing the evolution of the system in each abstract
state. We represent the dynamics using an affine model with an
interval error. Such a model can either be approximated using learning
methods or computed as a sound (over) approximation using reachability
set computation methods. Because we assume black box semantics, we
only present the former. Using regression analysis on the $start$ and
$end$ states of the witnessed trajectory segments between two cells,
we compute an approximate discrete map along with an error estimate.
Moreover, using the simulation function $\simulate$, additional
trajectory segments (or data) can be generated if required. The data
can be separated into a training set and testing set to compute the
map and the error respectively.  The latter case can be explored if
the symbolic dynamics of the system are known. A tool like
\flowstar~\cite{chen2013flow} can be used to find the reachable set
map.

Observe that $G^R$, a directed reachability graph, is rich enough to
search for concrete behaviors in the system. We call it a time
parameterized PWA relational abstraction. It can be interpreted as an
infinite state discrete transition system, and we can use
off-the-shelf bounded model checkers to find concrete violations of a
given safety property, and even other temporal properties.  We now
discuss the background required to present our ideas.

\section{Background}

In this section, we present the basics of relational abstractions and
fundamentals of regression analysis.

\subsection{Relational Abstraction}

For general hybrid dynamical systems, algorithmic ways for computing
analytical solutions do not exist. Discretization (along with suitable
abstractions) is often employed to transform the systems into a
discrete transition system. Relational abstractions is one such idea,
which abstracts continuous dynamics by discrete relations using
appropriate \textit{reachability invariants}. Both time independent
and time dependent relations have been proposed; the former captures
all reachable states over all time, whereas, the latter explicitly
includes time by relating reachable states to time. Thus it
can prove timing properties whereas, time independent relational
abstractions can not. In both cases, the resulting abstractions can be
interpreted as discrete transition systems and can be analyzed using
model checkers. We briefly discuss these two types of abstractions.

\mypara{Timeless Relational Abstraction}

Relational abstractions for hybrid systems were proposed in
\cite{Sankaranarayanan+Tiwari/2011/Relational}. For a hybrid automaton
model, they can summarize the continuous dynamics of each mode using a
binary relation over continuous states. The resulting relations are
timeless (independent) and hence valid for all time as long as
the mode invariant is satisfied. The relations take the general form
of $R(\x,\x') \bowtie 0$, where $\bowtie$ represents one of the
relational operators $=, \ge, \le, <, >$. For example, an abstraction
that captures the monotonicity with respect to time for the
differential equation $\dot{x} = 2$ is $x' > x$. The abstraction
capturing the relation between the set of ODEs: $\dot{x} = 2$,
$\dot{y} = 5$, is $5(x' - x) = 2(y' - y)$.

Such relation summaries, discretize the continuous evolution of a
given system into an infinite state transition system, in which safety
properties can be verified using $k$-induction, or falsified using
bounded model checkers. The relationalization procedure involves
finding suitable invariants in a chosen abstract domain, such as
affine abstractions, eigen abstractions or box abstractions
\cite{Sankaranarayanan+Tiwari/2011/Relational}.  For this, different
techniques like template based invariant generation
\cite{Gulwani+Tiwari/2008/Constraint,
Colon+Sankaranarayanan+Sipma/03/Linear} can be used.

\mypara{Timed and Time-Aware Relational Abstractions}

As the above discussed relations are timeless, we cannot reason over
the timing properties of the original system. Moreover, they are not
suitable for time-triggered systems. Timed relational
abstractions~\cite{zutshi2012timed} and time-aware relational
abstractions~\cite{mover2013time} were proposed to overcome these
shortcomings by explicitly including time in the relations. In
addition, both can be more precise than their timeless counterpart,
but unlike it, assume continuous affine dynamics.

To analyze time-triggered systems like an SDCS modeled by an affine
hybrid automata~\footnote{An affine hybrid automata is restricted to
the ODEs, resets and guards being affine.}, timed relational
abstractions can be computed directly from the solutions of the
underlying affine ODEs.  Recall that the solution of an affine ODE
$\dot{\x} = A\x$ can be represented using the matrix exponential as
$\x(t) = e^{tA}\x(0)$. This gives us the timed relation $\x' =
e^{tA}\x$.  Clearly, the relation is non-linear with respect to time.
However, for the case of SDCS with a fixed sampling time period, we
obtain linear relations in $\x$.  We demonstrated the usefulness of
timed relational abstractions for the case of linear systems
in \cite{zutshi2012timed}, and now incorporate the idea to discretize
black box dynamical systems.

Similar to timeless relational abstractions, time-aware relational
abstractions \cite{mover2013time} construct binary relations between
the current state $\x$ of the system and any future reachable state
$\x'$. The difference lies with the latter also constructing relations
between the current time $t$ and any future time $t'$. This is
achieved by a case by case analysis of the eigen structure of the
matrix $A$ (of an affine ODE). Separate abstractions are used for the
case of linear systems with constant rate, real eigen values and complex eigenvalues.

\subsection{Learning Dynamics using Simple Linear Regression}

\mypara{Regression Analysis}

In statistics, regression is the problem of finding a
\textit{predictor}, which can suitably predict the relationship
between the given set of observed input $\x$ and output $\y$ vectors.
In other words, assuming that $\y$ depends on $\x$, regression
strategies find either a parameterized or a non-parameterized
prediction function to explain the dependence. We now discuss simple
linear regression, which is parametric in nature and searches for an
affine predictor. It is also called \textit{ordinary least squares}.

\mypara{Ordinary Least Squares (OLS)}

Let the data set be comprised of $N$ input and output pairs $(\x,
y)$, where $\x\in\reals^n$ and $y\in\reals$. If $N>n$, which is
the case in the current context of \textit{finding the best fit}, the
problem is over-determined; there are more equations than
unknowns. Hence, a single affine function cannot be found which
satisfies the equation $\forall i\in\{1\ldots N\}. y_i = \vec{a}^T\x_i + b$. Instead, we
need to find the `best' choice for $\vec{a}$ and $b$. This is formally
defined using a loss function. For the case of simple linear
regression or OLS, the loss function is the sum of squares of the
errors in prediction.  The task is then to determine the vector of
coefficients $\vec{a}$ and an offset constant $b$, such that the least
square error of the affine predictor is minimized for the given data
set.
\begin{equation}
    \operatornamewithlimits{argmin}_{\vec{a}, b}\displaystyle\sum_{i=1}^{N}{\left(\y_i - (\vec{a}^T\x_i + b)\right)^2}
\label{reg}
\end{equation}
To ease the presentation, we can rewrite the above as a homogeneous
expression by augmenting $\x$ by $\hat{\x} = \begin{bmatrix}\x \\ 1\end{bmatrix}$ and replacing $\vec{a}$ and $b$ by the vector
$\vec{ab} = \begin{bmatrix}\vec{a} \\ b \end{bmatrix}$. The equation~\ref{reg} now becomes
\[
    \operatornamewithlimits{argmin}_{\vec{ab}}\displaystyle\sum_{i=1}^{N}{(\y_i - \vec{ab}^T \hat{\x}_i)^2}
\]
The solution of OLS can be analytically computed as
\[ Ab = (X^TX)^{-1}X^T\y\]
where $X$ is the matrix representing the horizontal stacking of all
$\hat{\x}$.
The details can be found in several texts on learning and statistics
\cite{friedman2001elements}.

Given a time invariant dynamical system $\x' = \simulate(\x, \Delta)$,
where $\x\in\reals^n$ we can use OLS to approximate its trajectories
at fixed time step $\Delta$ by a discrete map $\x' = A\x + \vb$, where
$A = \begin{bmatrix}\vec{a}_1^T \\ \vdots \\ \vec{a}_n^T \end{bmatrix}$ and $\vb =
\begin{bmatrix}b_1 \\ \vdots \\ b_n \end{bmatrix}$. This map is a
    global relational model for the system. The data set for the
    regression is a set of pairs $\setof{(\x,\x') | \x' =
    \simulate(\x, \Delta)}$ and can be generated on demand. Another
    set can be generated to compute the error $\delta$ as an interval
    of vectors, where each element is an interval $\delta_i \in
    [\delta_{min}, \delta_{max}]$.

    Affine maps are poor
    approximations for arbitrary non-linear functions. Hence, we use a
    collection of affine maps to approximate the local behaviors (in
    state-space) of the system $\simulate$. This results in a piece
    wise approximation, as described in the next section.

\subsection{Piecewise Affine (PWA) Transition System}

We define the PWA transition system as a transition system
(ref.~\secref{DiscSys}) $\rho:\tupleof{\pLocs,\pVars,\scrT, \pLocI,\Theta}$
where each transition $\tau \in \scrT$, is associated with a
transition relation of the form
\[
    \rho_{\tau}(\pVar,\pVar') \subseteq \setof{(\pVar,f_{\tau}(\pVar)) \;|\; g_\tau(\pVar) \land g'_\tau(f_{\tau}(\pVar))}
\]
where $f_{\tau}$ is an affine map relating the states and $g_{\tau}$,
$g'_{\tau}$ are affine guards (pre and post conditions) on the states.
The assertion on the initial states $\Theta$ is also affine.

We can now use a PWA transition system $\rho$ to approximate the
behavior of a hybrid dynamical system $\System$ defined by a simulator
$\simulate_{\System}$~\footnote{Due to $f_\tau$ being restricted to an affine
form, we can only approximate general hybrid systems.}. Each
transition of $\rho$ represents a discrete time step $\Delta$. Let the
state-space of $\System$ be given by $\x\in\HybridStates^n$, then the
states of $\rho$ are also given by $\x\in\HybridStates^n$. The affine guard
predicates $g_\tau$ are given by a conjunction of hyper-planes and
hence can be represented in matrix form as a polyhedron (defined by
$m$ constraints) in the state-space
\[
g:C\x - \vd \le 0
\]
where $C$ is an $m \times n$ matrix and $\vd$ is a vector of length
$m$. The affine maps $f_\tau$ can be represented in matrix form as
$f:\x \mapsto A\x+\vb$. To make $f_\tau$ `richer', we incorporate
an error term $\delta$ which is defined by a vector of intervals
$[\delta_{\min},\delta_{\max}]$. Hence,
\[
f:\x \mapsto A\x + \vb + \delta
\]
defines a set valued relation of the form $R:\setof{(\x,\x') \;|\; \x'
\in f_\tau(\x)}$.

% Given a state vector $\x\in\reals^n$, a guarded affine map
% $\gm:(\guard, \amap)$ defines the discretized consecution rule as a
% pair of an affine guard predicate $\guard:C\x - \vd \le 0$ and an affine map
% $\amap:A\x+\vb$, where $A \in \reals^{n \times n}$, $C \in
% \reals^{m \times n}$ are matrices and $\vb \in \reals^n$, $\vd
% \in \reals^m$ are vectors. A guarded affine map is satisfied if its
% guard is satisfied.

We now formalize the PWA affine transition system for a dynamical system.
\begin{definition}[PWA Transition System]

Given a hybrid dynamical system over a state-space
$\HybridStates^n$, a PWA transition system $\rho$ is given by the
tuple $\tupleof{\pLocs,\x,\scrT,\pLocI,\Theta}$, where $\tau \in
\scrT$ are affine transition relations and $\Theta$ is an affine
predicate over $\x$ and the states $\x\in\HybridStates^n$. The
transition relation is then defined by $\scrT$ with $n$ transition
relations as follows

\begin{equation}
    \scrT = \left\{
        \begin{array}{ll}
            g_1(\x) \land g'_1(\x') \implies \x' \in f_1(\x) \\
            \ldots\\
            g_n(\x) \land g'_n(\x') \implies \x' \in f_n(\x) \\
        \end{array}
    \right.
\end{equation}

\end{definition}

where $f_i:\x \mapsto A_i\x + \vb_i + \delta_i$,
$g_i(\x):C_i\x-\vd_i\le0$, $g'_i(\x'):C'_i\x'-\vd'_i\le0$ and $\x'$
denotes the next state of the system.

A PWA model is \textit{deterministic}, iff for every state
$\x\in\HybridStates^n$, a unique guarded affine map is satisfied and
all errors $\delta$ are singletons. However, in practice such a case
rarely exists. A PWA model will be usually non-deterministic, both due
to multiple choice of transition relations and the reachable states at
the $k^{th}$ time step (or after $k\Delta$ units of time) being a set.
Abusing the notation, we denote the set of states reachable in a
single step in $\rho$ by $\x' = \rho(\x)$.

A PWA transition system is \textit{complete}, iff for every state
$\x\in\HybridStates^n$, there exists at least one satisfied
transition relation $\scrT$. If this is not the case, the system can
deadlock, with no further executions. This usually results from
incorrect modeling of SDCS, and from here on, we do not consider such
cases.

% Given a dynamical system over a continuous state-space
% $\ContStates\in\reals^n$, a PWA model is a map $\ContStates \mapsto
% \setof{\gm_1, \gm_2, \ldots, \gm_n}$ from the state-space of the
% dynamical system to a finite set of guarded affine maps
% $\setof{\gm_1, \gm_2, \ldots, \gm_n}$. It defines the consecution
% rule by the affine map of the satisfied $\gm$.

% \begin{equation}
%     \pwa = \left\{
%         \begin{array}{ll}
%             \gm_1: \amap_1(\x) &\text{if}\; \guard_1(\x) \\
%             \gm_2: \amap_2(\x) &\text{if}\; \guard_2(\x) \\
%             \ldots & \ldots\\
%             \gm_n: \amap_n(\x) &\text{if}\; \guard_n(\x) \\
%         \end{array}
%     \right.
% \end{equation}

% \end{definition}

% where $\amap_i = A_i\x + \vb_i$, and $\guard_i(\x) \equiv C_i\x-\vd_i\le0$.
% Abusing the notation, we denote the next state computed by the PWA
% model as $\x' = \pwa(\x)$. A PWA model is \textit{deterministic}, iff
% for every state $\x\in\ContStates$, a unique guarded affine map is
% satisfied. A PWA model is \textit{complete}, iff for every state
% $\x\in\ContStates$, there exists at least one satisfied guarded affine
% map $\gm$.

>>>>

We further our exploration of trajectory segment based methods; and
explore an alternative approach to overcome the explosion in abstract
states. Instead of selectively refining the abstraction, we compute a
model of the black box system and use bounded model checking to find a
concrete counter-example in the model. Due to modeling errors, this
might not be reproducible in the original black-box system. We then
use this counter-example to guide the search towards a counter-example
in the original system.

@@

% Towards the end of the chapter, we discuss extensions to data driven
% approaches (instead of simulator driven), where instead of a
% simulator, a fixed set of data is provided as the behavioral
% description of the system. Using a combination of relational modeling,
% and program analysis we outline the future work for falsifying
% properties of SDCS.
